Untitled

### 2022-12-20 Figuring_Q3.Rmd
Working on figuring out Q3 (using a equal to current proposal value)

---
title: "Untitled"
output: html_document
date: "2022-12-14"
---
```{r}
# Part b
calcDirichletPDF <- function(x,a){
  #Calculating density of x under Dirichlet distribution with parameters specified by a
  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
  #calculating log of density
  log.density <- log(density)
  #returning a numeric vector
  return(c(log.density))
}

@@

>>>>

---
title: "Untitled"
output: html_document
date: "2022-12-14"
---
```{r}
# Part b
calcDirichletPDF <- function(x,a){

  #Calculating density of x under Dirichlet distribution with parameters specified by a
  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
  #calculating log of density
  log.density <- log(density)
  #returning a numeric vector
  return(c(log.density))
}

# x <- PT[i,,j-1]
# a <- 20*PT[i,,j-1]

#Initial values
X <- read.table("q3Weather.txt",header = F)
X <- X$V1
initPT <- matrix(1/3,nrow=3,ncol=3)
initStateProbs <- c(1/3,1/3,1/3)
n <- 100000
burnin <- 10000

#initializing algorithm
PT <- array(NA,dim=c(3,3,n+burnin+1))
PT[,,1] <- initStateProbs
proposal <- matrix(NA,ncol=3,nrow=3)

#Starting loop 
for(j in 2:(n+burnin+1)){

#Generating proposal value for each row of transition matrix (each weather state p_{s},p_{c},p_{r})
for (i in 1:length(initStateProbs)){
  proposal[i,] <- simDirichletRV(20*PT[i,,j-1])
}

#Generating posterior using equation (12) for proposal (new) and current value of PT
posterior.cur <- posterior.new <- NULL
for (i in 1:length(init.state)){
  posterior.cur[i] <- PT[init.state[i],trans.state[i],j-1]
  posterior.new[i] <- proposal[init.state[i],trans.state[i]]
}
# Calculating density of posterior using equation 12
posterior.density.cur <- prod(posterior.cur)*initStateProbs[X[1]]
posterior.density.new <-prod(posterior.new)*initStateProbs[X[1]]

# Calculating density using proposal density for each row (p_{x})
proposal.density.cur <- proposal.density.new <- NULL
for (i in 1:length(initStateProbs)){
proposal.density.new[i] <- calcDirichletPDF(proposal[i,],20*PT[i,,j-1])
proposal.density.cur[i] <- calcDirichletPDF(PT[i,,j-1],20*PT[i,,j-1])
}
#How do we combine the 3 vectors to compare with PT??? - do we just multiple to get joint probability density
proposal.density.cur <- exp(sum(proposal.density.cur))
proposal.density.new <- exp(sum(proposal.density.new))

#accept-reject
alpha <- (posterior.density.new*proposal.density.cur)/(posterior.density.cur*proposal.density.new)

if (runif(1) <= min(1,alpha) ){
  PT[,,j] <- proposal
}else{
  PT[,,j] <- PT[,,j-1]
}

}

######### Testing matrox selection
# PT[,,1] <- initPT <- matrix(c(1/2,0.25,0.25,1/3,1/6,0.5,1/8,1/4,5/8),nrow=3,ncol=3,byrow = T)
# PT[2,,1]

########
```

```{r}

# Part b
calcDirichletPDF <- function(x,a){

  #Calculating density of x under Dirichlet distribution with parameters specified by a
  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
  #calculating log of density
  log.density <- log(density)
  #returning a numeric vector
  return(c(log.density))
}

#Initial values
X <- read.table("q3Weather.txt",header = F)
X <- X$V1
initPT <- matrix(1/3,nrow=3,ncol=3)
initStateProbs <- c(1/3,1/3,1/3)
n <- 100000
burnin <- 10000

q3.mcmc <- function(X,initStateProbs,initPT,burnin,n){
#Converting X into a numeric vector to access entries of PT
X <- ifelse(X=="s",1,(ifelse(X=="c",2,3)))
init.state <- X[1:(length(X)-1)]
trans.state <- X[2:length(X)]

#initializing algorithm
PT <- array(NA,dim=c(3,3,n+burnin+1))
PT[,,1] <- initStateProbs
proposal <- matrix(NA,ncol=3,nrow=3)

#Starting loop 
for(j in 2:(n+burnin+1)){

  #Generating proposal value for each row of transition matrix (each weather state p_{s},p_{c},p_{r})
  for (i in 1:length(initStateProbs)){
    proposal[i,] <- simDirichletRV(20*PT[i,,j-1])
  }

  #Generating posterior using equation (12) for proposal (new) and current value of PT
  posterior.cur <- posterior.new <- NULL
  for (i in 1:length(init.state)){
    posterior.cur[i] <- PT[init.state[i],trans.state[i],j-1]
    posterior.new[i] <- proposal[init.state[i],trans.state[i]]
  }
  # Calculating density of posterior using equation 12
  posterior.density.cur <- prod(posterior.cur)*initStateProbs[X[1]]
  posterior.density.new <-prod(posterior.new)*initStateProbs[X[1]]

  # Calculating density using proposal density for each row (p_{x})
  proposal.density.cur <- proposal.density.new <- NULL
  for (i in 1:length(initStateProbs)){
    proposal.density.new[i] <- calcDirichletPDF(proposal[i,],20*PT[i,,j-1])
    proposal.density.cur[i] <- calcDirichletPDF(PT[i,,j-1],20*PT[i,,j-1])
  }
  #How do we combine the 3 vectors to compare with PT??? - do we just multiple to get joint probability density
  proposal.density.cur <- exp(sum(proposal.density.cur))
  proposal.density.new <- exp(sum(proposal.density.new))

  #accept-reject
  alpha <- (posterior.density.new*proposal.density.cur)/(posterior.density.cur*proposal.density.new)

  if (runif(1) <= min(1,alpha) ){
    PT[,,j] <- proposal
  }else{
    PT[,,j] <- PT[,,j-1]
  }

}
#Removing burnin samples
return(PT[,,(burnin+1):(n+burnin+1)])
}

set.seed(2022)

PT.sample <- q3.mcmc(X,initStateProbs,initPT,burnin,n)

table(X)

PT.expected <- matrix(c(mean(PT.sample[1,1,]),mean(PT.sample[1,2,]),mean(PT.sample[1,3,]),
                        mean(PT.sample[2,1,]),mean(PT.sample[2,2,]),mean(PT.sample[2,3,]),
                        mean(PT.sample[3,1,]),mean(PT.sample[3,2,]),mean(PT.sample[3,3,])),ncol=3,nrow=3,byrow = T)
round(PT.expected,3)
```

```{r}
# Part b
calcDirichletPDF <- function(x,a){
  #Calculating density of x under Dirichlet distribution with parameters specified by a
  density <- (gamma(sum(a))/prod(gamma(a)))*prod(x^(a-1))
  #calculating log of density
  log.density <- log(density)
  #returning a numeric vector
  return(c(log.density))
}

@@

```{r}
## Trying again monday 19th dec

###Testing 
#Initial values
#finding prob of sequence under current pt
X <- read.table("q3Weather.txt",header = F)
X <- X$V1
initPT <- matrix(1/3,nrow=3,ncol=3)
initStateProbs <- c(1/3,1/3,1/3)
n <- 100000
burnin <- 500

      #Generating posterior using equation (12) for proposal and old value of PT
      actualp.prev <- actualp.prop <- NULL
      for (i in 1:length(init.state)){
        actualp.prev[i] <- V[init.state[i],trans.state[i],j-1]
        actualp.prop[i] <- proposal[init.state[i],trans.state[i]]
      }
      #density has some constant as proportional but due to fraction constant cancels out so dont
      #need to worry about it
      true.density.prev <- prod(actualp.prev)*initStateProbs[X[1]]
      true.density.prop <-prod(actualp.prop)*initStateProbs[X[1]]

      #Finding density of propoisal using Diriclet(20*prev.value)
      prop.dens.prop <- prop.dens.prev <- NULL
       for (i in 1:length(initStateProbs)){
        prop.dens.prop[i] <- calcDirichletPDF(proposal[i,],20*V[i,,j-1])
        prop.dens.prev[i] <- calcDirichletPDF(20*V[i,,j-1],20*V[i,,j-1])

       }
      total.prop.dens.prop <- exp(sum(prop.dens.prop))
      total.prop.dens.prev <- exp(sum(prop.dens.prev))

      # alpha <- (true.density.prop*prop.density.prev)/(true.density.prev*prop.density.prop)
      alpha <-  (true.density.prop/true.density.prev)*(total.prop.dens.prev/total.prop.dens.prop)

      if(min(1,alpha)>=runif(1)){
        V[,,j] <- proposal
      }else{
        V[,,j] <- V[,,j-1]
      }
}
```
